Geometrization leftovers

### 2013-07-30 base change.tex
Moved base change discussion to a new file

% !TEX encoding = UTF-8 Unicode
\documentclass[11pt]{amsart}
\title[Geometration of characters of tori]{Geometrization of characters of tori over non-Archimedean local fields}
\usepackage[british]{babel}
\usepackage{datetime}
\date{\today}
\author{Clifton Cunningham}
\address{University of Calgary}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Pacific Institute for the Mathematical Sciences at the University of Calgary}
\email{roed.math@gmail.com}
\usepackage[utf8]{inputenc}
\renewcommand{\baselinestretch}{1.2}
\usepackage[notcite,color]{showkeys}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage[shortalphabetic]{amsrefs}
\usepackage[alphabetic]{amsrefs}
\renewcommand\MR{\relax}
\usepackage{xypic}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{yfonts}
\newcommand{\mathswab}[1]{\operatorname{\textswab{#1}}}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
\usepackage{manfnt}

@@

\newcommand{\FF}{{\mathbb{F}}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\NN}{{\mathbb{N}}}
\newcommand{\CC}{{\mathbb{C}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\RR}{{\mathbb{R}}}
\newcommand{\EE}{\mathbb{\bar Q}_\ell}
\newcommand{\OK}{\mathcal{O}_K}
\newcommand{\pK}{\mathfrak{p}_K}
\newcommand{\OL}{\mathcal{O}_L}
\newcommand{\Zp}{\mathbb{Z}_p}
\newcommand{\Qp}{\mathbb{Q}_p}
%\newcommand{\Fp}{\mathbb{F}_p}
\newcommand{\bFq}{\bar{k}}
\newcommand{\Fq}{k}
\newcommand{\Fqm}{k_m}
%\newcommand{\bFp}{{\mathbb{\bar F}_p}}

@@

\newcommand{\cdef}[1]{{{\color{cyan}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\TL}{\underline{T_L}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\GC}{{\mathcal{G\hskip-0.8pt C}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CS_0(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}

@@

\begin{abstract}
  We study the category of character sheaves on smooth commutative
  group schemes $G$ locally of finite type over finite fields
  $\Fq$. Assuming that the geometric component group of $G$
  is finitely-generated, we show that the group of isomorphism classes
  of character sheaves on $G$ is canonically isomorphic to the group
  of characters of $G(\Fq)$. We use this result to exhibit an
  isomorphism between admissible characters of $T(K)$ and isomorphism
  classes of character sheaves on the Greenberg transform of the Néron
  model of $T$, where $K$ is any non-Archimdean local field and $T$ is
  any algebraic torus over $K$.
\end{abstract}

\maketitle

\section{Introduction}

%\input{introduction}
%In his celebrated thesis, John Tate writes 
%\begin{quotation}
%    {\it Concerning the characters of $k^*$ [the
%    multiplicative group of a $\mathfrak{p}$-adic field], the
%    situation is different from that of $k^+$.  First of all, we are
%    interested in continuous multiplicative maps $\alpha \mapsto
%    c(\alpha)$ of $k^*$ into the complex numbers, not only in the
%    bounded ones, and shall call such a map a quasi-character,
%    reserving the word ``character'' for the conventional character of
%    absolute value $1$. Secondly, we shall find no model for the group
%    of quasi-characters, or even for the group of characters, though
%    such a model would be of the utmost importance.}
%  \cite{Tate:thesis}*{\S 2.3}
%\end{quotation}
%\noindent
In this paper we show that the monidal category of character sheaves
on the Greenberg transform of the locally finite type Neron model of
$T$ provides an elegant, canonical model for the group of
(quasi-)characters of $T(K)$, where $K$ is any non-Archimedean local
field and $T$ is any algebraic torus over $K$.

Since the character sheaves that appear in this paper are not quite an
instance of those defined by George Luzstig, who coined the term, we
begin by explaining what we mean by `character sheaves' here. 
Let $G$ be a smooth commutative group scheme locally of finite
type over $\Fq$; $G$ need not be reductive, connected, nor of
finite type over $\Fq$. Fix a prime $\ell$ that does not divide
$q$. By a {character sheaf} on $G$ we mean a triple $\cs{L}\ceq
(\gcs{L},\mu,F)$ where $\gcs{L}$ is a constructible $\ell$-adic sheaf
$\gcs{L}$ on the \'etale site of $\bG \ceq G \times_{\Spec\Fq}
\Spec{\bFq}$, locally constant of rank-$1$ on every connected
component of $\bG$, equipped with an isomorphism $\mu: m^*
\gcs{L} \to \LxL$, where $m : \GxG \to G$ is
the group operation for $G$, and an isomorphism $F : \Frob_{G}^*
\gcs{L} \to \gcs{L}$, where $\Frob_{G}$ is the Frobenius automorphism
on $\bG$, satisfying natural compatibility conditions with $\mu$;
these compatibility conditions are explained below.

Character sheaves on $G$ form a category, denoted by $\CS(G)$. In
fact, $\CS(G)$ is a rigid symmetric monoidal category, so isomorphism
classes of character sheaves on $G$ form a group, denoted by
$\CSiso{G}$; this is the Picard group of the category of
character sheaves on $G$. Our application of character sheaves to
characters of algebraic tori over non-Archimedean local fields rests
on the Picard group of character sheaves on $G$, for all smooth
commutative group schemes $G$, locally of finite type over $\Fq$, 
for which the geometric component group is finitely generated.

We now say a few words about the group schemes 
that appear in our geometrization of admissible
characters of algebraic tori over local fields. Let $K$ be a
non-Archimedean local field with residual field $\Fq$; we place no
restrictions on the characteristic of $K$ -- in particular, it may be
$0$ or not. Write $\OK$ for ring of integers of $K$ and $\pK$ for the
maximal ideal in $\OK$.  Let $T$ be an algebraic torus defined over
$K$. Let $\TT$ be a N\'eron model for $T$; recall that this is a
smooth group scheme locally of finite type over $\OK$. For each
positive integer $d$, let $\GN{T}_d$ be the Greenberg transform of
$\TT \otimes \OK/\pK^d$; the commutative group scheme $\GN{T}_d$
plays the r\^{o}le of $G$, above.

The main result of this paper is a canonical isomorphism
\[
\CSiso{\GN{T}_d} 
\iso
\Hom_{< d}(T(K),\EEx)
\] 
between the Picard group of the category of character sheaves on
$\GN{T}_d$ and the group of characters $T(K) \to \EEx$ of
depth at most $d$.  Since these isomorphisms are compatible with
respect to $d$, we also .....  \David{I'll do this when we're closer to being finished.}  
%In particular, by
%specializing to the case $T= \Gm{K}$ and by
%choosing an isomorphism $\EE \approx \CC$ (unnatural though it may
%be), we arrive at a model for the group of quasi-characters of
%$K^\times$: the monoidal category of character sheaves on the
%Greenberg transform of the Neron model of
%$\Gm{K}$.

Structure of the paper:

Thanks to:

\tableofcontents

\section{The Picard group of character sheaves}\label{section:category}

%\input{category}
%\subsection*{Smooth commutative group schemes}

Throughout this paper, $G$ stands for a commutative group scheme,
smooth and locally of finite type over a finite field $\Fq$; we write
$m : \GxG\to G$ for the multiplication morphism. We fix an
algebraic closure $\bFq$ of $\Fq$ and write $\bG $ for the
commutative group scheme, smooth and locally of finite type over
$\bFq$ obtained by base change; the multiplication morphism for
$\bG$ will be denoted by $\bm$. Let $\Frob_{G}$
(resp. $\Frob_{\GxxG}$) be the Frobenius automorphism of
$\bG$ (resp. $\bGxG$).
\Clifton{I'll insert a reference to SGA where Frobenius is defined for
  a large class of schemes over finite fields.}

%\subsection*{The category of character sheaves}

The category of character sheaves on $G$, denoted by \cdef{$\CS(G)$},
is defined as follows.  A \cdef{character sheaf} on $G$ is a triple
$\cs{L}\ceq (\gcs{L},\mu,F)$ where:
\begin{enumerate}
\labitem{(CS.0)}{CS.0} $\gcs{L}$ is a constructible $\ell$-adic
  sheaf on (the \'etale site of) $\bG$, locally constant of rank
  $1$ on connected components of $\bG$;

\labitem{(CS.1)}{CS.1} $\mu: \bm^* \gcs{L} \to \LxL$
  is an isomorphism of sheaves on $\bGxG$ such that

  \[
  \begin{tikzcd}[row sep=40]
  \bm_3^*\gcs{L} \arrow{rr}{(\bm \tight{1}{\times}{2} \id)^*\mu} \arrow[swap]{d}{(\id \tight{2}{\times}{1} \bm)^*\mu}
    &&  \bm^*\gcs{L} \boxtimes \gcs{L} \dar{\mu \tight{0}{\boxtimes}{1} \id} \\
    \gcs{L} \boxtimes \bm^* \gcs{L} \arrow{rr}{\id \boxtimes \mu}
    &&  \gcs{L} \tight{0}{\boxtimes}{0} \LxL\\
  \end{tikzcd}
  \]
  commutes, where $m_3 \ceq m\circ (m\tight{1}{\times}{2}\id) = m\circ (\id\tight{2}{\times}{1} m)$;

\labitem{(CS.2)}{CS.2} $F : \Frob_{G}^* \gcs{L} \to \gcs{L}$ is an
  isomorphism of sheaves on $\bG$ compatible with $\mu$ in the sense that

  \[
  \begin{tikzcd}
  \Frob_{\GxxG}^* \bm^* \gcs{L} \arrow{rr}{\Frob_{\GxxG}^*\mu}
    && \Frob_{\GxxG}^*(\LxL)\\
    \arrow[equal]{u} \bm^*  \Frob_{G}^* \gcs{L} \arrow[swap]{d}{\bm^* F}
    && \Frob_{G}^*\gcs{L}\boxtimes \Frob_{G}^*\gcs{L} \dar{F\boxtimes F} \arrow[equal]{u} \\
    \bm^*\gcs{L} \arrow{rr}{\mu}
    && \LxL\\
  \end{tikzcd}
  \]
  commutes.
\end{enumerate}
Morphisms in $\CS(G)$ are defined in the natural way:
\begin{enumerate}
\labitem{(CS.3)}{CS.3} if $\cs{L} = (\gcs{L},\mu,F)$ and
  $\cs{L'} = (\gcs{L'},\mu',F')$ are character sheaves on $G$ then
  $\Hom(\cs{L},\cs{L'})$ is the set of morphisms $\varphi : \gcs{L} \to \gcs{L'}$
  of constructible $\ell$-adic sheaves on $\bG$ such that
  \[
  \begin{tikzcd}[column sep=40]
  \Frob_G^* \gcs{L} \rar{\Frob_G^* \varphi} \arrow[swap]{d}{F} & \Frob_G^* \gcs{L'} \dar{F'}
  & \arrow[draw=none]{d}[pos=.4,description]{\text{\normalsize{and}}} & \arrow[swap]{d}{\mu} m^* \gcs{L} \rar{m^* \varphi} & m^* \gcs{L'} \dar{\mu'} \\
  \gcs{L} \rar{\varphi} & \gcs{L'}
  & {} & \LxL \rar{\tight{1}{\varphi\boxtimes \varphi}{1}} & \gcs{L'} \tight{0}{\boxtimes}{0} \gcs{L'}
  \end{tikzcd}
  \]
  both commute.
\end{enumerate}

%\subsection*{The Picard group of character sheaves}

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', F\otimes F')$
with duals formed by applying the functor $\sheafHom(\ - \ ,(\EE)_{\bG})$ to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group} of character
sheaves on $G$; {\it cf.} {\tt http://ncatlab.org/nlab/show/Picard+group}.
\Clifton{Replace ncatlab with other references.}

\begin{remark} \label{rem:CS-action}
If $W$ is a group equipped with a homomorphism $\alpha : W \to \Gal(\bFq/\Fq)$
then an \emph{action} of $W$ on $\gcs{L}$ is a collection of isomorphisms
\[
\gamma_w : \alpha(w)^* \gcs{L} \to \gcs{L}
\]
so that $\gamma_{uw} = \gamma_w \circ \alpha(w)^*\gamma_u$. \David{make sure I have the order right and add a reference}
Specifying an $F$ as in \ref{CS.2} is the same as giving an action of the Weil group $\Weil{k}$
on $\gcs{L}$ compatible with $\mu$.
\end{remark}

%\subsection*{Trace of Frobenius}\label{sec:Frob}

Let $\cs{L}$ be a character sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob_G$ and
therefore an automorphism $F_{\bg}$ of $\gcs{L}_{\bg}$. Note
that $\Aut_{\EE}(\gcs{L}_{\bg}) \iso \EEx$, canonically (this
follows from second condition appearing in the definition of character
sheaves). The isomorphism $\mu : \bm^* \gcs{L} \to \LxL$
guarantees that the function $t^\Frob_\cs{L} : G(\Fq)\to \EEx$,
defined by $g \mapsto F_{\bg}$, is a group homomorphism;
we refer to $t^\Frob_\cs{L}$ as the \cdef{trace of Frobenius of $\cs{L}$}.
\Clifton{There is no actual trace involved, since $\gcs{L}$ is a rank
  $1$ local system. More generally, trace is needed.}

It is clear that $t^\Frob_{\cs{L}\otimes \cs{L'}} = t^\Frob_\cs{L} \cdot t^\Frob_\cs{L'}$
(pointwise multiplication of functions) and that $t^\Frob_\cs{L} = t^\Frob_\cs{L'}$
if $\cs{L} \iso \cs{L'}$ in $\CS(G)$. In this way we obtain a group homomorphism
\[
\cdef{t_{G}} : \CSiso{G} \longrightarrow \Hom_\text{grp}(G(\Fq),\EEx),
\qquad \text{defined by}\qquad \cs{L} \mapsto t^\Frob_\cs{L},
\]
from the Picard group of character sheaves on $G$ to the character group of $G(\Fq)$.

Our main result on the geometrization of admissible characters of
$p$-adic tori [insert internal reference here] follows from a general
fact established later in this paper: if the geometric component group
of $G$ is finitely-generated then the trace of Frobenius defines an
isomorphism from the Picard group of character sheaves $G$ to the
group of characters of $G(\Fq)$ (here, as throughout the paper, $G$ is
a commutative group scheme, smooth and locally of finite type over $\Fq$).

%\subsection*{Pull-back of character sheaves}\label{subsection:pullback}

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,F) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \CS(G) \to \CS(H)$ such that
  \[
  \begin{tikzcd}[row sep=40]
   \CSiso{G} \rar{f^*} \arrow[swap]{d}{t_{G}} & \CSiso{H} \dar{t_H} \\
  \Hom_\text{grp}(G(\Fq),\EEx) \rar & \Hom_\text{grp}(H(\Fq),\EEx)
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

\begin{proof}
  Let $\cs{L}$ be a character sheaf on $G$. We start by showing that
  $\bf^*\gcs{L}$ is locally constant on connected components of
  $\bH$. Let $c_j :\bH^j \hookrightarrow \bH$ be any
  connected component and let $i$ be the image of $j$ under the group
  homomorphism $\pi_0(\bG) \to \pi_0(\bH)$ obtained by
  applying the functor $\pi_0$ to $\bf$; see \cite{vdG&M}*{III, 3.28}, for example.
  \Clifton{Replace with a better reference, eventually.}
  Write $c_i : \bG^i \hookrightarrow \bG$ for
  the inclusion of $\bG^i$ and $\bf^j : \bH^j \to \bG^i$
  for the restriction of $f$ to $\bH^j$.  Since $\cs{L}$
  is a character sheaf, $\gcs{L}$ is locally constant on the connect
  components of $\bG$; thus, there is a finite etale covering
  $a_i : X_i \to \bG^i$ such that $a_i^* (\gcs{L}\vert_{\bG^i})$
  is constant. Let $b_j : Y_j \to \bH^j$ be the
  pull-back of $a_i$ along $\bf^j$; let $g_j : Y_j \to X_i$ be
  the map opposite $\bf^j$. Then $b_j$ is a finite etale covering
  of $\bH^j$ and
  \[
  b_j^* \left( (\bf^* \gcs{L})\vert_{\bH^j}\right)
  = (\bf\circ c_j \circ b_j)^*\gcs{L}
  = (c_i\circ a_i\circ g_j)^*\gcs{L} = g_j^* a_i^* (\gcs{L}\vert_{\bG^i}).
  \]

  Since $a_i^* (\gcs{L}\vert_{\bG^i})$ is a constant sheaf and since the pull-back
  of a constant sheaf (along $g_j$, in this case) is a constant sheaf,
  it follows that $b_j^* \left( (\bf^*\gcs{L})\vert_{\bH^j}\right)$
  is a constant sheaf. Thus, $\bf^*\gcs{L}$ is locally constant on
  $\bH^j$. Since $(\bf^*\gcs{L})_{\bg} = \gcs{L}_{f(\bg)}$ for every
  geometric point ${\bg}$ on $\bG$, and since $\gcs{L}$ has
  rank $1$ on connected components of $\bG$, it follows that
  $f^*\gcs{L}$ also has rank $1$ on connected components of $\bG$.
  This completes the proof that $\bf^*\gcs{L}$ satisfies condition \ref{CS.0}.

  To see that $(\bfxf)^* \gcs{L}$ satisfies
  condition \ref{CS.1} simply apply the functor $(\bfxf)^*$
  to \ref{CS.1} for $\cs{L}$ and use the canonical isomorphism
  $(\bfxf)^*(\LxL) \iso \bf^*\gcs{L} \tight{-3}{\boxtimes}{-3} \bf^*\gcs{L}$
  (a few times). Likewise, to see that $\bf^*\gcs{L}$ satisfies condition
  \ref{CS.2}, apply the same functor to \ref{CS.2} for $\gcs{L}$.
  Then use the fact that $f$ is defined over $\Fq$, so
  $(\bfxf)^*\Frob_{\GxxG}^* \iso \Frob_{\GxxG} (\bfxf)^*$,
  and the fact that $f$ is a morphism of group schemes, so
  $(\bfxf)^* \bm^*\iso \bm^* \bf^*$
  (isomorphisms of functors on constructible sheaves).

  To see that $(\gcs{L},\mu,F) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a functor $f^* : \CS(G) \to \CS(H)$ one applies the functors $\bf^*$ and
  $\bf^*\tight{1}{\times}{1}\bf^*$ to \ref{CS.3} and argue as
  above. It follows from the fact that sheaf hom commutes with these
  functors that $f^* : \CS(G) \to \CS(H)$ is a monoidal functor.

  The fact that the functor $f^* : \CS(G) \to \CS(H)$ commutes with
  the trace of Frobenius, in the sense above, follows immediately from
  the definitions; see \cite{Laumon}*{1.1.1.2}, where the ambient
  hypothesis that $X$ is of finite type over $\Fq$ can be replaced by
  the hypothesis that $X$ is locally of finite type over $\Fq$.

  Finally, the fact that $(f\circ g)^* = g^* \circ f^*$ follows from the analogous
  statements about pullbacks of sheaves and maps of sheaves.
  \David{Is this explanation of $(f\circ g)^* = g^* \circ f^*$ sufficient?}
\end{proof}

Consider the short exact sequence in the category of smooth group
schemes, locally of finite type over $\Fq$, defining the component
group scheme for $G$:
\begin{equation}\label{eq:pi-2}
\begin{tikzcd}
0 \rar & G^0 \rar & G \rar & \pi_0(G) \rar & 0.
\end{tikzcd}
\end{equation}
Since $\pi_0(G)$ is an Abelian group scheme, smooth (in fact, etale) and locally of finite type over $\Fq$, 
it follows from Proposition~\ref{prop:pullback} that \eqref{eq:pi-1} defines a sequence of functors
\[
\begin{tikzcd}
\CS(0) \rar & \CS(\pi_0(G)) \rar & \CS(G) \rar & \CS(G^0) \rar & \CS(0)
\end{tikzcd}
\]
which induces, upon passage to Picard groups, a sequence of group homomorphisms 
\begin{equation}\label{eq:pi-1}
\begin{tikzcd}
\CSiso{0} \rar & \CSiso{\pi_0(G)} \rar & \CSiso{G} \rar & \CSiso{G^0} \rar & \CSiso{0}.
\end{tikzcd}
\end{equation}
It is easy to see that $\CSiso{0}$ is trivial. In this paper we will find $\CSiso{G^0}$ and $\CSiso{\pi_0(G)}$ under the hypothesis that $\pi_0(G)$ is geometrically finitely generated; the snake lemma then allows us to find $\CSiso{G}$.

\section{Bounded character sheaves}

In this section we describe full subcategory of $\CS(G)$, 
called the category of bounded character sheaves on $G$ and denoted by $\CS_b(G)$, 
which will illuminate the nature of $\CS(G)$ and will also play a role in the arguments later in this paper.

Consider the category $\GC(G)$ of finite etale covers of $G$ that are also morphisms of abelian group schemes.
Recall that we fixed an algebraic closure $\bFq$ of $\Fq$ long ago; 
let ${\bar e} : \Spec{\bFq} \to G$ be the geometric point on $G$ obtained by composing $\Spec{\bFq} \to \Spec{\Fq}$ with the identity $e : \Spec{\Fq} \to G$ for $G$.
Let $\ker_{\bar e}$ be the set-valued functor on $\GC(G)$
defined by pull-back along the geometric point ${\bar e} : \Spec{\bFq} \to G$ 
followed by forgetting the structure sheaf and topology; 
thus, if $f : H\to G$ is a finite etale morphism of group schemes, then $\ker_{\bar e}(f)$ is the set underlying 
$\ker{\bar f} = H_{\bar e}$.
Let $\pi_1^\text{gp}(G)$ be the automorphism group of the functor $\ker_{\bar e}$. 
We will see that $\pi_1^\text{gp}(G)$ is a profinite abelian group.
A \cdef{bounded character sheaf on $G$} is a continuous irreducible representation of $\pi_1^\text{gp}(G)$. 
Let $\CS_b(G)$ be the monoidal category of irreducible continuous $\ell$-adic representations of $\pi_1^\text{gp}(G)$.

\begin{proposition}
 $G\mapsto \pi_1^\text{gp}(G)$ is a functor from group schemes over $\Fq$ to profinite abelian groups, and .....
\end{proposition}

\begin{proof}
The category $\GC(G)$, together with the set-valued functor $\ker_{\bar e}$ satisfy conditions (G1) through (G6) appearing in \cite[Exposé V, \S 4, p. 118-119]{SGA1a}.
\end{proof}

\section{Connected abelian algebraic groups}

%\input{connected-1}
% !TEX encoding = UTF-8 Unicode
%In this section we review a fundamental property of character
%sheaves on geometrically connected, commutative algebraic groups over $\Fq$.

\begin{proposition}\label{prop:connected-1}
  If $G$ is a geometrically connected commutative algebraic group over $\Fq$ then
  \[
  t_{G} : \CSiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx),
  \]
  is an isomorphism of groups.
\end{proposition}

\begin{proof}
  We will prove this proposition by reducing it to the functions--sheaves dictionary
  \cite{Laumon}*{1.1}. The main point of this proof will therefore be to show that,
  {\it since $G$ is geometrically connected}, every character sheaf on $G$ descends to
  a local system on $G$ satisfying the analogue of \ref{CS.1} on $G$.

  First, observe that the forgetful functor $(\gcs{L},\mu,F) \mapsto (\gcs{L},F)$
  sends character sheaves on $G$ to Weil local systems on
  $G$ \cite{Deligne:Weil2}*{1.1.10} (by a Weil local system on $G$ we
  mean a Weil sheaf on $G$ which is a local system on $\bG$).
  While is it not true that all Weil local systems on $G$
  descend to sheaves on $G$, those that appear in the image of the
  forgetful functor from character sheaves do.

  To see why this is so, consider the category \cdef{$\CS_0(G)$} of
  pairs $(\cs{L}_0,\mu_0)$ where $\cs{L}_0$ is a local system on $G$
  and $\mu_0: m^*\cs{L}_0 \to \cs{L}_0 \tight{0}{\boxtimes}{0} \cs{L}_0$ satisfies
  condition \ref{CS.1} adapted to sheaves on $G$; maps in $\CS_0(G)$
  are defined by the second part of condition \ref{CS.2}, {\it mutatis mutandis}.
  \David{This section should be modified once we're done with the section on bounded character sheaves}
  Let $b : \bG \to G$ be base change
  (pull-back of $G\to \Spec{\Fq}$ along $\Spec{\bFq} \to \Spec{\Fq}$).
  The proof of Proposition~\ref{prop:pullback} adapts
  (simplifies, in fact) to show that $b^* : \CS_0(G) \to \CS(G)$ is a
  monoidal functor. This functor is fully faithful; see
  \cite{SGA7.2}*{Expos\'e XIII} and \cite{BBD}*{Prop. 5.2.1}.  We will
  show that $b^*$ is essentially surjective.

  Let $g$ be any $\Fq$-rational point on $G$ (such as the identity for
  $G$) and let $\bg$ be the geometric point on $G$ lying above
  $g$ (we fixed $\Fq \hookrightarrow \bFq$ long ago). Since $G$ is
  geometrically connected, the geometric point $\bg$ determines
  an equivalence between the category of Weil local systems on $G$ and
  $\ell$-adic representations of $\W(G,\bg)$
  \cite{Deligne:Weil2}*{1.1.12}. Now let $(\gcs{L},\mu,F)$ be a
  character sheaf on $G$. Then the isomorphism class of the Weil local
  system $(\gcs{L},F)$ corresponds to an $\ell$-adic character of the
  Weil group $\W(G, \bg)$. The $\Fq$-rational point $g$ under
  the geometric point $\bg$ determines a splitting
  $\Weil{\Fq}\to \W(G,\bg)$ of the exact sequence
  \[
  \begin{tikzcd}
  1 \rar & \pi_1(\bG, \bg) \rar &  \W(G,\bg) \rar
  & \ar[swap,bend right, dotted]{l}{g} \Weil{\Fq} \rar & 1
  \end{tikzcd}
  \]
  In this way the Weil local system $(\gcs{L},F)$ determines an $\ell$-adic character
  \[
  t_{(\gcs{L},F)}(g) : \Weil{\Fq} \to \EEx \qquad \text{defined by}\qquad \Frob \mapsto t^\Frob_{(\gcs{L},F)}(g)
  \]
  of the Weil group $\Weil{\Fq}$; the image of $\Frob$ under this
  character is $t^\Frob_\cs{L}(g)$, the trace of Frobenius of $\cs{L}$ at $g$.

>>>>

% !TEX encoding = UTF-8 Unicode
\documentclass[11pt]{amsart}
\title[Geometrization leftovers]{Geometrization leftovers}
\usepackage[british]{babel}
\usepackage{datetime}
\date{\today}
\author{Clifton Cunningham}
\address{University of Calgary}
\email{cunning@math.ucalgary.ca}
\author{David Roe}
\address{Pacific Institute for the Mathematical Sciences at the University of Calgary}
\email{roed.math@gmail.com}
\usepackage[utf8]{inputenc}
\renewcommand{\baselinestretch}{1.2}
\usepackage[notcite,color]{showkeys}
\usepackage{hyperref}
\usepackage{geometry}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage[shortalphabetic]{amsrefs}
\usepackage[alphabetic]{amsrefs}
\renewcommand\MR{\relax}
\usepackage{xypic}
\usepackage{textcomp}
\usepackage{mathrsfs}
\usepackage{yfonts}
\newcommand{\mathswab}[1]{\operatorname{\textswab{#1}}}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,calc,matrix}
\usepackage{tikz-cd}
\usepackage{manfnt}

@@

\newcommand{\FF}{{\mathbb{F}}}
\newcommand{\ZZ}{{\mathbb{Z}}}
\newcommand{\NN}{{\mathbb{N}}}
\newcommand{\CC}{{\mathbb{C}}}
\newcommand{\QQ}{{\mathbb{Q}}}
\newcommand{\RR}{{\mathbb{R}}}
\newcommand{\EE}{\mathbb{\bar Q}_\ell}
\newcommand{\OK}{\mathcal{O}_K}
\newcommand{\pK}{\mathfrak{p}_K}
\newcommand{\OL}{\mathcal{O}_L}
\newcommand{\OO}[1]{\mathcal{O}_{#1}}
\newcommand{\Zp}{\mathbb{Z}_p}
\newcommand{\Qp}{\mathbb{Q}_p}
%\newcommand{\Fp}{\mathbb{F}_p}
\newcommand{\bFq}{\bar{k}}
\newcommand{\Fq}{k}
\newcommand{\Fqm}{k_m}
%\newcommand{\bFp}{{\mathbb{\bar F}_p}}

@@

\newcommand{\cdef}[1]{{{\color{cyan}#1}\index{#1}}} 
\newcommand{\sheafHom}{{\mathscr{H}\hskip-4pt{\it o}\hskip-2pt{\it m}}}
\newcommand{\abs}[1]{{\vert #1 \vert}}
\newcommand{\ceq}{{\, :=\, }}
\newcommand{\tq}{{\ \vert\ }}
\newcommand{\iso}{{\ \cong\ }}
\newcommand{\obj}{{\text{obj}\, }}
\newcommand{\Gm}[1]{\mathbb{G}_{\hskip-1pt\textbf{m},#1}}
\newcommand{\GN}[1]{\mathswab{#1}}
\newcommand{\bGN}[1]{{\bar{\mathswab{#1}}}}
\newcommand{\TT}{\underline{T}}
\newcommand{\TL}{\underline{T_L}}
\newcommand{\invlim}[1]{\lim\limits_{\overleftarrow{#1}}}
\newcommand{\dirlim}[1]{\lim\limits_{\overrightarrow{#1}}}
\newcommand{\limit}[1]{\mathop{\textsc{lim}}\limits_{#1}}
\newcommand{\colimit}[1]{\mathop{\textsc{colim}}\limits_{#1}}
\newcommand{\cs}[1]{{\mathcal{#1}}}
\newcommand{\gcs}[1]{{\mathcal{\bar #1}}}
\newcommand{\dualgcs}[1]{\gcs{#1}^\dagger}
\newcommand{\dualcs}[1]{\cs{#1}^\dagger}
\newcommand{\CS}{{\mathcal{C\hskip-1.8pt S}}}
\newcommand{\GC}{{\mathcal{G\hskip-0.8pt C}}}
\newcommand{\CSiso}[1]{\CS(#1)_{/\textit{iso}}}
\newcommand{\CSbiso}[1]{\CS_0(#1)_{/\text{iso}}}
\newcommand{\CE}{{\mathcal{C\hskip-1.8pt E}}}
\newcommand{\Lgroup}[1]{{\,^L\hskip-1pt{#1}}}
\newcommand{\dualgroup}[1]{{\check{#1}}}
\newcommand{\Lang}{{\operatorname{Lang}}}
\newcommand{\image}{{\operatorname{im}}}
\newcommand{\Weil}[1]{\mathcal{W}_{#1}}
\newcommand{\level}[1]{(\mathrm{level} N)}

@@

\section{Base change over the residue field}
